---
title: "Lemon sharks ecology Bimini"
author: "Sprinkles"
date: "7/27/2021"
output:
  html_document: default
  pdf_document: default---
---

```{r clear_memory}
rm(list = ls())
```

## Data loading

Set the working directory depending on the user

```{r setup}
print(Sys.info()["nodename"])
if (grepl("aurits",Sys.info()["nodename"])){
  data.dir<-"~/Documents/Science/Projects/Rob Bullock - Bimini/data/"}
```

Load libraries

```{r load_library,result='hide',message=F}
library(dplyr)
library(devtools)
#install_git('https://gitlab.com/bartk/move.git') #Installs 'move' development version
library(move)
library(ggplot2)
library(maptools)
library(circular)
library(ggmap)
library(mapproj)
library(knitr)
```

Load data set

```{r read_files}
DET<-read.csv(file.path(data.dir,"/TRACKS1.csv"))
```

Check data directory and create it if not present

```{r data directory}
#set path to general folder where output files will be written
out.dir<-"dBBMM ASCII/"

if(!dir.exists(data.dir)) stop(paste0("Directory not found: \n",
                                      "       ", gsub("\\.", getwd(), data.dir)))
if(!dir.exists(out.dir)) dir.create(paste(data.dir,out.dir,sep=""))
```

Define some global variables

```{r global_vars}
sep=","
dec="."
dat.TZ="US/Eastern"
```

Explore the data set

```{r expl}
names(DET)
str(DET)
dim(DET)
head(DET)
table(DET$Tidal.Phase)
table(DET$Shark,DET$Tidal.Phase)
table(DET$Shark)
```

Housekeeping and tidy up

```{r hk}
DET$Datetime  <- as.POSIXct(DET$Datetime,format="%m/%d/%y %H:%M",tz=dat.TZ)
DET$Tag       <- as.numeric(DET$Shark)
DET$Lat       <- DET$N
DET$Lon       <- DET$W
DET$T.Ph      <- DET$Tidal.Phase

DET <- dplyr::select(DET,Datetime,Shark,Tag,T.Ph,Lat,Lon)
DET <- arrange(DET,Tag,Datetime)
```

Convert lonlat to UTM

```{r coord_UTM}
# First convert coordinate sets to SpatialPoints and project
cord.dec = SpatialPoints(cbind(DET$Lon, DET$Lat), proj4string=CRS("+proj=longlat"))

# Transform to UTM by setting the EPSG to 32617 for WGS 84, UTM zone 17, northern hemisphere. This is where Bimini is located.
cord.UTM <- as.data.frame(spTransform(cord.dec, CRS("+init=epsg:32617")))
colnames(cord.UTM) <- c("NewEastingUTM", "NewNorthingUTM")
DET <- cbind(DET,cord.UTM)
```

Construct movement models per individual. 

Some notes regarding the 'move package and construction of movement models:
Several arguments need to be used to run the model.
1. Window size: corresponds with number of locations and moves along a given trajectory to estimate the MA parameter within defined subsections of the path. This increases the ability to detect breakpoints where changes in behaviour occur. The window size should relate to what kind of behaviours the model is desired to identify e.g., a window size of 23 means the sliding window is moved every 23 locations or every 23 hours (has to do with sampling interval)
2. Margin: motion variance based on only the middle section of the trajector; the ends of the movement trajectory where no changes are allowed because at some stage you want to have a few locations to base your estimation of the variance on and how many locations in either side of the window we use for this, is called the margin. Smaller values for window size and margin is expected to give a higher frequency of behavioural changes; make these large for looking at migrations.
3. The raster dictates the grid cell size for the UD to be calculated per grid cell per individual. Create the raster of certain size that matches with coordinates used to make the move object
4. Extent: is incorporated if there are animal locations that border the edges of the raster.

A dBBMM ia nor run if total detections of individual < window size (default value, 31). Below code checks and filters individuals with insufficient data. Below we set window size arbitrarily to 23

```{r filter_data}
check1 <- DET %>%
  group_by(Tag) %>%
  summarise(relocations=length(Datetime))
check1
check2 <- filter(check1,relocations>23)
check2
  
if(length(check1$Tag)!=length(check2$Tag)){
  DET <- semi_join(DET,check2)
  check1 <- DET %>%
    group_by(Tag) %>%
    summarise(relocations=length(Datetime))
  check1
  check2 <- filter(check1,relocations>23)
  length(check1$Tag)==length(check2$Tag)
}
```

In the next R code we create per individual a move object, project it, contruct a dBBMM, calculate the volume area within the 50% and 95% contours and finally save the outcome as ASCII for plotting in an external GIS software.

```{r construct_dBMMMM}
bb <- list()
bb.list <- list()

DET<-DET[-411,]

# Loop through all unique tags
counter<-0
for (i in unique(DET$Tag)) {
  
  # Print individual
    counter<-counter+1
  
  # Filter individual data
    DET.i<-DET[DET$Tag==i,]
    
  # create move object
    move.i <- move(x=DET.i$Lon,y=DET.i$Lat,time=DET.i$Datetime,
                     proj=CRS("+proj=longlat +datum=WGS84"),
                     data=DET.i,animal=DET.i$Tag,sensor="VR2W")
  
  # Check the currect projection
    proj4string(move.i)
    head(move.i)
    
  # Incorporate uncertainty in the model by including a location error. From communication with Rob, the gps location of a shark is estimated, from the boat coordinates, bearing and distance estimate, to be 1 m.
    move.i$LocationError = 1 # 1 m: gps loc of shark inferred from boat coord + bearing + distance estimate
    head(move.i)
    
  # Convert projection to Azimuthal Equi-Distance projection (aeqd)
    r.i <- spTransform(move.i,center=TRUE)
  
  # Make sure it changed correctly
    proj4string(r.i)
  
  # Calculate time lag between consecutive detections. Obviously there is no time lag for the first detection, so we need to tell R this. Knowing time lags in acoustic telemetry is important, as the motion variance is based on the time difference b/w detections at consecutive locations i.e. larger time lag creates a wider bridge where the animal could've been and therefore potentially inflates the motion variance. Below code deals with this issue by identifying the location and number of detections with large time gaps. Below we will use an arbitrary value of 2 h.
    TimeDiff <- timeLag(r.i, units="hours")
    r.i$TimeDiff <- append(0,TimeDiff)
    long <- which(r.i$TimeDiff>2)
    long
    length(long) 
  
  # Make a raster for the UD to plot into. Start with UTM. These coordinates need to be big enough to cover your data. May need to expand x and y ranges (i.e. the "e" variable below) if you encounter errors when constructing the DBBMM in the next code chunk. NOTE: "xUTM" grid should be larger than the "x" i.e. lower minimums and higher. The resolution determines the grid size. The variable 'e' influences the extent. Finer resolution/grid size (in m) i.e. lower value means longer computation time. May need/want to play with this value too.
    e <- 80*1000
    xUTM.i <- raster(xmn=min(DET.i$NewEastingUTM)-e, xmx=max(DET.i$NewEastingUTM)+e, ymn=min(DET.i$NewNorthingUTM)-e, 
                    ymx=max(DET.i$NewNorthingUTM)+e,crs =CRS("+proj=utm +zone=17 +datum=WGS84"), 
                    resolution=50)
  
  # We now need to reproject this into aeqd. Make a dummy object to get the correct projection.
    x.i <- raster(xmn=min(DET.i$NewEastingUTM),xmx=max(DET.i$NewEastingUTM),ymn=min(DET.i$NewNorthingUTM),
                  ymx=max(DET.i$NewNorthingUTM),crs=CRS("+proj=utm +zone=17 +datum=WGS84"))
    proj4string(x.i)
  
  # Use that to reproject UTM to AEQD
  # ?projectExtent # project the values of a Raster object to a new Raster object with another projection (CRS), i.e. projectExtent(from, to, ...)) 
    newTemplate <- projectExtent(xUTM.i,proj4string(x.i))
    newTemplate # get the number of pixels in newTemplate and write that number into the next function. This is an empty raster.
  
  # Give newTemplate some values. Make Rep equal to the ncell dimension
    ones = rep(1, ncell(newTemplate))
    xAEQD.i <- setValues(newTemplate, ones)
    xAEQD.i
    
    # Reproject the move object r into AEQD
    rNew.i <- spTransform(r.i, proj4string(xAEQD.i))
    rNew.i
  
  # We need to make sure that the projection of the move object is in the same format as our raster. Check below.
    proj4string(xAEQD.i)
    proj4string(rNew.i)
  
  # Below we exclude the data points that are 'long' i.e. create a large time gap.
    bursted <- burst(rNew.i, c('normal','long')[1+(timeLag(rNew.i, units = 'hours')>2)]) 
  # There are 2 types of burst: "normal" and "long". You can select for these in the dbbmm by selecting the factor level in the burstType command.
  
  # Construct the model. The time.step should reflect the ping frequency of the tag (in minutes)
    bursted_dbbmm <- brownian.bridge.dyn(bursted, burstType='normal', raster = xAEQD.i, 
                                         location.error = "LocationError", time.step = 5, ext = 3,
                                         window.size = 23)
  
  # Re-standardize (Dr. Kranstauber's trouble shooting solution). Occurring errors are "due to limits of accuracy during calculations. Given the error is really small (<.000001 %) I would not worry to much and re-standardize". Then aggregate UD segments.
    tmp <- calc(bursted_dbbmm,sum)
    bb <- new(".UD",tmp/sum(values(tmp)))
  
 # Export aggregated individual UDs to as ascii files for further exploration/spatial analysis in GIS software. Needed for when the aim is to plot population-level and normalized UDs per species.
    asc <- writeRaster(bb, paste(data.dir,out.dir, filename=i,".asc",sep=""), 
                     format="ascii", datatype="FLT4S", bylayer=T, overwrite=T)
    
  # Calculate volume area (m^2) within 50% (core) and 95% (general use) contours
    area.50 <- sum(values(getVolumeUD(bb)<=.50))
    area.95 <- sum(values(getVolumeUD(bb)<=.95))
    
  # Combine in single df
    area.ct <- data.frame(core.use = area.50, general.use = area.95)
    
  # Add Shark and Tag id
    area.ct$Shark<-unique(DET$Shark)[i]
    area.ct$Tag<-i
    
    
  # Put in list
    bb.list[[counter]]<-area.ct
    
    bb.list
    
    gc()
    
}

# Put everything in a data.frame
  md<-bind_rows(bb.list, .id = "column_label")
  
write.table(md, file=paste(data.dir,"Lemon_VolumeArea.txt",sep=""),row.names=F,sep=",",dec=".")

  
```

Now that we have constructed individual movement models for complete movement trajectories, below we will do the same, but now we will discern different tidal phases when constructing the models.

```{r dir}
#set path to general folders where output files will be written
out.dir_h<-"dBBMM ASCII/1"
if(!dir.exists(paste(data.dir,out.dir_h,sep=""))) stop(paste0("Directory not found: \n",
                                      "       ", gsub("\\.", getwd(), paste(data.dir,out.dir_h,sep=""))))
if(!dir.exists(out.dir_h)) dir.create(paste(data.dir,out.dir_h,sep=""))

out.dir_l<-"dBBMM ASCII/2"
if(!dir.exists(paste(data.dir,out.dir_l,sep=""))) stop(paste0("Directory not found: \n",
                                      "       ", gsub("\\.", getwd(), paste(data.dir,out.dir_l,sep=""))))
if(!dir.exists(out.dir_l)) dir.create(paste(data.dir,out.dir_l,sep=""))

out.dir_m<-"dBBMM ASCII/3"
if(!dir.exists(paste(data.dir,out.dir_m,sep=""))) stop(paste0("Directory not found: \n",
                                      "       ", gsub("\\.", getwd(), paste(data.dir,out.dir_m,sep=""))))
if(!dir.exists(out.dir_m)) dir.create(paste(data.dir,out.dir_m,sep=""))
```

```{r construct_tidal_dbbmm}
# Transform Tide class to numeric
DET$Tide<-as.numeric(DET$T.Ph)

# Loop through the different tidal phases
for (i in unique(DET$Tide)) {
  
  # Filter individual tide data
    DET.t<-DET[DET$Tide==i,]
    
  # remove sharks with insufficient movement data (identical to code L113-L128)
    check1 <- DET.t %>%
      group_by(Tag) %>%
      summarise(relocations=length(Datetime))
    check1
    check2 <- filter(check1,relocations>23)
    check2
  
    if(length(check1$Tag)!=length(check2$Tag)){
      DET.t <- semi_join(DET.t,check2)
      check1 <- DET.t %>%
        group_by(Tag) %>%
        summarise(relocations=length(Datetime))
      check1
      check2 <- filter(check1,relocations>23)
      length(check1$Tag)==length(check2$Tag)
    }
    
  # Loop through all unique tags
    
    bb.list<-list()
    counter<-0
    for (j in unique(DET.t$Tag)) {
      
      # Print individual
        counter<-counter+1
  
      # Filter individual data
        DET.i<-DET.t[DET.t$Tag==j,]
    
      # create move object
        move.i <- move(x=DET.i$Lon,y=DET.i$Lat,time=DET.i$Datetime,
                         proj=CRS("+proj=longlat +datum=WGS84"),
                         data=DET.i,animal=DET.i$Tag,sensor="VR2W")
  
      # Check the currect projection
        proj4string(move.i)
        head(move.i)
    
      # Incorporate uncertainty in the model by including a location error. From communication with Rob, the gps location of a shark is estimated, from the boat coordinates, bearing and distance estimate, to be 1 m.
        move.i$LocationError = 1 # 1 m: gps loc of shark inferred from boat coord + bearing + distance estimate
        head(move.i)
    
      # Convert projection to Azimuthal Equi-Distance projection (aeqd)
        r.i <- spTransform(move.i,center=TRUE)
  
      # Make sure it changed correctly
        proj4string(r.i)
  
      # Calculate time lag between consecutive detections. Obviously there is no time lag for the first detection, so we need to tell R this. Knowing time lags in acoustic telemetry is important, as the motion variance is based on the time difference b/w detections at consecutive locations i.e. larger time lag creates a wider bridge where the animal could've been and therefore potentially inflates the motion variance. Below code deals with this issue by identifying the location and number of detections with large time gaps. Below we will use an arbitrary value of 2 h.
        TimeDiff <- timeLag(r.i, units="hours")
        r.i$TimeDiff <- append(0,TimeDiff)
        long <- which(r.i$TimeDiff>2)
        long
        length(long) 
  
      # Make a raster for the UD to plot into. Start with UTM. These coordinates need to be big enough to cover your data. May need to expand x and y ranges if you encounter errors when constructing the DBBMM in the next code chunk. NOTE: "xUTM" grid should be larger than the "x" i.e. lower minimums and higher. The resolution determines the grid size. The variable 'e' influences the extent. Finer resolution/grid size (in m) i.e. lower value means longer computation time.
        e <- 80*1000
        xUTM.i <- raster(xmn=min(DET.i$NewEastingUTM)-e, xmx=max(DET.i$NewEastingUTM)+e, ymn=min(DET.i$NewNorthingUTM)-e, 
                        ymx=max(DET.i$NewNorthingUTM)+e,crs =CRS("+proj=utm +zone=17 +datum=WGS84"), 
                        resolution=50)
  
      # We now need to reproject this into aeqd. Make a dummy object to get the correct projection.
        x.i <- raster(xmn=min(DET.i$NewEastingUTM),xmx=max(DET.i$NewEastingUTM),ymn=min(DET.i$NewNorthingUTM),
                      ymx=max(DET.i$NewNorthingUTM),crs=CRS("+proj=utm +zone=17 +datum=WGS84"))
        proj4string(x.i)
  
      # Use that to reproject UTM to AEQD
      # ?projectExtent # project the values of a Raster object to a new Raster object with another projection (CRS), i.e.     projectExtent(from, to, ...)) 
        newTemplate <- projectExtent(xUTM.i,proj4string(x.i))
        newTemplate # get the number of pixels in newTemplate and write that number into the next function. This is an empty raster.
  
      # Give newTemplate some values. Make Rep equal to the ncell dimension
        ones = rep(1, ncell(newTemplate))
        xAEQD.i <- setValues(newTemplate, ones)
        xAEQD.i
    
        # Reproject the move object r into AEQD
        rNew.i <- spTransform(r.i, proj4string(xAEQD.i))
        rNew.i
  
      # We need to make sure that the projection of the move object is in the same format as our raster. Check below.
        proj4string(xAEQD.i)
        proj4string(rNew.i)
  
      # Below we exclude the data points that are 'long' i.e. create a large time gap.
        bursted <- burst(rNew.i, c('normal','long')[1+(timeLag(rNew.i, units = 'hours')>2)]) 
      # Ther are 2 types of burst: "normal" and "long". You can select for these in the dbbmm by selecting the factor level in the burstType command.
  
      # Construct the model. The time.step should reflect the ping frequency of the tag (in minutes)
        bursted_dbbmm <- brownian.bridge.dyn(bursted, burstType='normal', raster = xAEQD.i, 
                                             location.error = "LocationError", time.step = 5, ext = 3,
                                             window.size = 23)
  
      # Re-standardize (Dr. Kranstauber's trouble shooting solution). Occurring errors are "due to limits of accuracy during calculations. Given the error is really small (<.000001 %) I would not worry to much and re-standardize". Then aggregate UD segments.
        tmp <- calc(bursted_dbbmm,sum)
        bb <- new(".UD",tmp/sum(values(tmp)))
  
      # Export aggregated individual UDs to as ascii files for further exploration/spatial analysis in GIS software. Needed for when the aim is to plot population-level and normalized UDs per species.
       asc <- writeRaster(bb, paste(data.dir,out.dir, unique(DET$Tide)[i],"/", filename=j,".asc",sep=""), 
                     format="ascii", datatype="FLT4S", bylayer=T, overwrite=T)
    
      # Calculate volume area (m^2) within 50% (core) and 95% (general use) contours
        area.50 <- sum(values(getVolumeUD(bb)<=.50))
        area.95 <- sum(values(getVolumeUD(bb)<=.95))
    
      # Combine in single df
        area.ct <- data.frame(core.use = area.50, general.use = area.95)
    
     # Add Shark and Tag id
        area.ct$Shark<-unique(DET$Shark)[j]
        area.ct$Tag<-j
    
      # Put in list
        bb.list[[counter]]<-area.ct
    
        bb.list
    
        gc()
        
    }

  # Put everything in a data.frame
    md<-bind_rows(bb.list, .id = "column_label")
        
    assign(paste0("md",".",unique(DET$T.Ph)[i],sep=""), md)
    
    write.table(assign(paste0("md",".",unique(DET$T.Ph)[i],sep=""), md), x = file=paste(data.dir,"Lemon_VolumeArea_",unique(DET$T.Ph)[i],".txt",sep=""),row.names=F,sep=",",dec=".")

}s

```


